---
title: "Lab Section 3"
author: "Napon Jatusripitak"
date: "4/19/2019"
output: pdf_document
---
# Lab Section 3
## Dplyr

Provides useful functions for easy data manipulation:  
- ```filter()``` allows you to select specific observations  
- ```arrange()``` allows you to reorder your data  
- ```select()``` allows you to select data based on names  
- ```rename()``` allows you to rename variables  
- ```mutate()``` allows you to create new variables from current ones  
- ```summarize()``` condenses data values  
- ```sample_n()``` takes samples  

Here are some examples:
```{r message=FALSE, warning=FALSE, results="hide"}
# Load dplyr
library(dplyr)

# Examples
filter(mtcars, mpg < 20, cyl == 6) # filter(df, condition, condition)
arrange(mtcars, mpg) # arrange(matcars, desc(mpg)) returns the result in a descending order
select(mtcars, mpg, cyl) # select(df, variable1, variable2)
mutate(mtcars, kpg = mpg*1.609) # create a new variable inside the dataframe called kpg as a function of mpg
```

## Pipes
Consult this [guide](https://style.tidyverse.org/pipes.html) for additional info. Generally, we use pipes to organize our codes into readable, sequential chunks. Here's how I use them. Recall the ```starwars``` dataset.

- What is the average mass for blue-eyed characters that are taller than 160 but shorter than 200?

```{r message=FALSE, warning=FALSE, results="hide"}
# Load the packages & the data
library(dplyr)
library(broom)
data(starwars)
```

```{r}
### Without pipes
subset_starwars <- filter(starwars, eye_color == "blue", height > 160 & height < 200)
mean(subset_starwars$mass, na.rm=T)

### With pipes
starwars %>%
  filter(eye_color == "blue", height > 160 & height < 200) %>%
  summarize(Mean = mean(mass, na.rm=T))
```
- Calculate the number of characters of each species that have mass greater than 60
```{r}
### Without pipes
subset_starwars <- filter(starwars, mass > 60)
group <- group_by(subset_starwars, species)
result <- tally(group)
result

### With pipes
starwars %>%
  filter(mass > 60) %>%
  group_by(species) %>%
  tally()
```

Now it's your turn. Use ```%>%``` to complete the following steps.

- Calculate the average height by species (Hint: use ```group_by()``` and ```summarize()```)
```{r eval=FALSE, include=FALSE}
starwars %>%
  group_by(species) %>%
  summarize(Mean = mean(height))
```

- Calculate the mass to height ratio for all characters and rank them from high to low. (Hint: use ```mutate()``` to create a new column for mass to height ratio and use ```arrange()``` to rank the observations)
```{r eval=FALSE, include=FALSE}
starwars %>%
  mutate(mass_to_height = mass/height) %>%
  arrange(desc(mass_to_height)) %>%
  select(name, mass_to_height) #optional
```

- What is the coefficient on ```height``` when we regress ```mass``` on ```height```?
```{r eval=FALSE, include=FALSE}
height_coef <- lm(mass~height, data=starwars) %>% coef()
height_coef
```

- Construct a dataframe from the regression model's results (Hint: use ```tidy()``` from the ```broom``` library)
```{r eval=FALSE, include=FALSE}
lm(mass~height, data=starwars) %>% tidy()
```

## Regression & Classification
In this exercise, we will be predicting whether a given character in the ```starwars``` dataset is human or non-human, using mass and height as our predictors and OLS as our model.  

1. Import the dataset as ```starwars```. Drop any missing values in ```species```, ```mass``` and ```height```.
```{r eval=FALSE, include=FALSE}
data(starwars)
starwars %>% drop_na(species, mass, height) -> starwars # that arrow at the end assigns the result to starwars, essentially modifying the original dataset
```
2. Get rid of Jabba Desilijic Tiure. He's a little too heavy for our analysis!
```{r eval=FALSE, include=FALSE}
char_outlier <- select(filter(starwars, mass == max(starwars$mass)), name) # a roundabout way to do this. Get the charater with the maximum mass and remove him from the dataset.
starwars <- filter(starwars, name != as.character(char_outlier))
# starwars <- filter(starwars, name !="Jabba Desilijic Tiure") would also work
```
3. Estimate a regression in the form of $y = x_1 + x_2 + \epsilon$ and display the results. (Hint: create a binary response variable that takes the value of 1 if the character is a human and 0 if the character is not a human.)

```{r eval=FALSE, include=FALSE}
starwars$response <- ifelse(starwars$species == 'Human', 1, 0)
```

```{r eval=FALSE, include=FALSE}
ols <- lm(response ~ mass + height, data = starwars)
summary(ols)
```

4. Set our decision boundary. Let's say that $y \ge 0.5$ implies that the character is a human and $y < 0.5$ implies that the character is not a human. How good is our prediction? (Hint: create a confusion matrix and compute the accruracy/classification rate.)


```{r eval=FALSE, include=FALSE}
# predicted values
pred <- ifelse(predict(ols, type = "response") >= 0.5, "Human", "Non-Human")
real = ifelse(starwars$response == 1, "Human", "Non-Human")
# assess accuracy/performance
table(real, predicted = pred)
mean(real == pred)
```
5. Create a scatterplot of mass against height. Use color or shape to distinguish between human and non-human characters. Add a line where $\hat{y} = 0.5$ as a function of mass and height (Hint: use ```coef()```.)

```{r eval=FALSE, include=FALSE}
intercept <- (0.5 - coef(ols)[1]) / coef(ols)[3]
slope <- -1 * coef(ols)[2] / coef(ols)[3]
```


```{r eval=FALSE, include=FALSE}
library(ggplot2)
ggplot(starwars, aes(x = mass, y = height, color = real)) +
geom_point(size = 2) +
geom_abline(intercept = intercept, slope = slope, linetype = 'dashed') +
labs(color='Species')
```
6. Challenge: repeat the previous steps, using Logit instead of OLS.

